### 2022-11-29 report/article_2.tex
initalized latex template and wrote content

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.4 (15/5/16)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com) with extensive modifications by
% Vel (vel@LaTeXTemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[twoside,twocolumn]{article}

\usepackage{blindtext} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[english]{babel} % Language hyphenation and typographical rules

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text

\usepackage{enumitem} % Customized lists
\setlist[itemize]{noitemsep} % Make itemize lists more compact

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\roman{subsection}} % roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\usepackage{titling} % Customizing the title section

\usepackage{hyperref} % For hyperlinks in the PDF

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\setlength{\droptitle}{-4\baselineskip} % Move the title up

\pretitle{\begin{center}\Huge\bfseries} % Article title formatting
\posttitle{\end{center}} % Article title closing formatting
\title{A parralelized and distributed approach to implementing Conway's Game of Life} % Article title
\author{%
\textsc{Aaron Chan}\thanks{A thank you or further information} \\[1ex] % Your name
\normalsize University of bristol \\ % Your institution
\normalsize {ho21739@bristol.ac.uk} % Your email address
\and % Uncomment if 2 authors are required, duplicate these 4 lines if more
\textsc{Ferdinand Hubbard}\thanks{Corresponding author} \\[1ex] % Second author's name
\normalsize University of bristol \\ % Your institution
\normalsize {ej21378@bristol.ac.uk} % Your email address
}
\date{\today} % Leave empty to omit a date
\renewcommand{\maketitlehookd}{
\begin{abstract}
  In this report, we will explain our two solutions for implementing Conway's Game of Life in golang. The first solution is designed to be run on a
  single machine, and makes use of all the cpu threads available on a computer through the parralelization of the program. The second solution is designed
  to run on multiple machines, allowing for the work to be distributed accross multiple machines. This second solution uses aspects of the first solution
  within each worker.
\end{abstract}
}

%----------------------------------------------------------------------------------------

\begin{document}

% Print the title
\maketitle

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------
\section{Introduction to the parralelized solution}

% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
\begin{itemize}
  \item single threaded implemementation
  \item events passing (alive cells, turn completed)
  \item PGM image output implementation
  \item SDL implemementation (flipped cells events)
  \item using channels and goroutines to create multiple workers (parralelizing the solution)
  \item switching to a memory sharing solution instead of using channels (TODO?)
  \item implementing halo exchange
\end{itemize}
%------------------------------------------------

\section{benchmarking methodology}

\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
\blindtext % Dummy text
%------------------------------------------------

\section{parralelizing the program}

\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
\blindtext % Dummy text

%------------------------------------------------

\section{parralelizing benchmark and critical analysis}

\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
\blindtext % Dummy text

%------------------------------------------------

\section{memory sharing}

\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
\blindtext % Dummy text

%------------------------------------------------

\section{benchmarking and critical analysis of the memory sharing}

\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
\blindtext % Dummy text

%------------------------------------------------

\section{Implementing Halo Exchange}

\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
\blindtext % Dummy text

%------------------------------------------------

\section{benchmarking and critical analysis of the Halo Exchange}

\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
\blindtext % Dummy text

%------------------------------------------------

\section{benchmarking and critical analysis of enabling SDL}

\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
\blindtext % Dummy text

%------------------------------------------------

\section{comparing the parralel benchmarks}

Maecenas sed ultricies felis. Sed imperdiet dictum arcu a egestas. 
\begin{itemize}
\item Donec dolor arcu, rutrum id molestie in, viverra sed diam
\item Curabitur feugiat
\item turpis sed auctor facilisis
\item arcu eros accumsan lorem, at posuere mi diam sit amet tortor
\item Fusce fermentum, mi sit amet euismod rutrum
\item sem lorem molestie diam, iaculis aliquet sapien tortor non nisi
\item Pellentesque bibendum pretium aliquet
\end{itemize}
\blindtext % Dummy text

Text requiring further explanation\footnote{Example footnote}.

%------------------------------------------------

\section{Results}

\begin{table}
\caption{Example table}
\centering
\begin{tabular}{llr}
\toprule
\multicolumn{2}{c}{Name} \\
\cmidrule(r){1-2}
First name & Last Name & Grade \\
\midrule
John & Doe & $7.5$ \\
Richard & Miles & $2$ \\
\bottomrule
\end{tabular}
\end{table}

\blindtext % Dummy text

\begin{equation}
\label{eq:emc}
e = mc^2
\end{equation}

\blindtext % Dummy text

%------------------------------------------------

\section{Discussion}

\subsection{Areas for improvement}

A statement requiring citation \cite{Figueredo:2009dg}.
\blindtext % Dummy text

\subsection{Thanks}

\blindtext % Dummy text

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
Figueredo, A.~J. and Wolf, P. S.~A. (2009).
\newblock Assortative pairing and life history strategy - a cross-cultural
  study.
\newblock {\em Human Nature}, 20:317--330.

\end{thebibliography}

%----------------------------------------------------------------------------------------

\end{document}

### 2022-11-29 report/article_2.tex
more content

@@
 
-\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-\blindtext % Dummy text
+% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
+\begin{itemize}
+  \item use built in golang benchmark framework
+  \item run command ....
+  \item output to txt
+  \item run python script to plot graphs
+  \item barchart of time taken against threads used
+  \item runs benchmarks on all pgm images
+  \item repeats benchmarks until results fall within a certain standard deviation (TBC)
+\end{itemize}
 %------------------------------------------------
@@
 
-\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-\blindtext % Dummy text
-
+% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
+\begin{itemize}
+  \item show workers pseudocode
+  \item show handler pseudocode
+  \item flowchart/ diagram of how channel communication workers
+  \item simple flowchart of how the whole program works currently
+\end{itemize}
 %------------------------------------------------
@@
 
-\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-\blindtext % Dummy text
-
+% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
+\begin{itemize}
+  \item show graph for 5120x5120.gpm for single implementation vs parralel implementation
+  \item describe the differences between each graph
+  \item explain the differences by referencing the flowchart/ how the code worker
+  \item explain the shape of the graph (why it resembles a 1/x graph). e.g. bottlenecks in channel communication...
+\end{itemize}
 %------------------------------------------------
@@
 
-\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-\blindtext % Dummy text
-
+% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
+THIS SECTION WILL BE USED IF WE DECIDE TO IMPLEMENT MEMORY SHARING OVER CHANNEL COMMUNICATION
 %------------------------------------------------
@@
 
-\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-\blindtext % Dummy text
-
+% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
+TO BE IMPLEMENTED
 %------------------------------------------------
@@
 \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-\blindtext % Dummy text
-
+TO BE IMPLEMENTED
 %------------------------------------------------
@@
 
-\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-\blindtext % Dummy text
-
+% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
+\begin{itemize}
+  \item show graphs and describe shape and differences of with vs without SDL
+  \item explain how SDL works
+  \item explain how this causes bottlenecks
+\end{itemize}
 %------------------------------------------------

### 2022-12-02 report/article_2.tex
fixed misspelling, para 1

@@
 
-\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
+\usepackage[hmarginratio=5:5,top=32mm,columnsep=20pt]{geometry} % Document margins
 \usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
@@
 \posttitle{\end{center}} % Article title closing formatting
-\title{A parralelized and distributed approach to implementing Conway's Game of Life} % Article title
+\title{A parallelised and distributed approach to implementing Conway's Game of Life} % Article title
 \author{%
-\textsc{Aaron Chan}\thanks{A thank you or further information} \\[1ex] % Your name
-\normalsize University of bristol \\ % Your institution
+\textsc{Aaron Chan}\\[1ex] % Your name
+\normalsize University of Bristol \\ % Your institution
 \normalsize {ho21739@bristol.ac.uk} % Your email address
 \and % Uncomment if 2 authors are required, duplicate these 4 lines if more
-\textsc{Ferdinand Hubbard}\thanks{Corresponding author} \\[1ex] % Second author's name
-\normalsize University of bristol \\ % Your institution
+\textsc{Ferdinand Hubbard}\\[1ex] % Second author's name
+\normalsize University of Bristol \\ % Your institution
 \normalsize {ej21378@bristol.ac.uk} % Your email address
@@
 \begin{abstract}
-  In this report, we will explain our two solutions for implementing Conway's Game of Life in golang. The first solution is designed to be run on a
-  single machine, and makes use of all the cpu threads available on a computer through the parralelization of the program. The second solution is designed
-  to run on multiple machines, allowing for the work to be distributed accross multiple machines. This second solution uses aspects of the first solution
-  within each worker.
+  We intend to explain our two solutions for implementing Conway's Game of Life in Golang, namely the Parallel and Distributed versions.
+  Within this report, we will discuss the impacts of thread usage in both our implementations, as well as the effect of distributed computation as opposed to single-machine computation.
+In essence, we found that execution time was sped up as we increased thread count, reduced branch instructions and implemented a halo exchange mechanism.
 \end{abstract}
@@
 %----------------------------------------------------------------------------------------
-\section{Introduction to the parralelized solution}
+\section{Introduction to the parallelised solution}
 
-% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
+As an initial attempt, we began by implementing a single-threaded Game of Life (GoL) Engine, expanding upon this work with parallelisation by virtue of concurrent go-routine workers.
+Recognising that branch instructions throttled efficiency of \texttt{getNeighbourCount}, we removed if-statements, introduced bit-masking and assignment which drastically improving runtime.
+
 \begin{itemize}
-  \item single threaded implemementation
+  \item single threaded implementation
   \item events passing (alive cells, turn completed)
   \item PGM image output implementation
-  \item SDL implemementation (flipped cells events)
-  \item using channels and goroutines to create multiple workers (parralelizing the solution)
+  \item SDL implementation (flipped cells events)
+  \item using channels and goroutines to create multiple workers (parallelising the solution)
   \item switching to a memory sharing solution instead of using channels (TODO?)
@@
 
-\section{parralelizing the program}
+\section{parralelising the program}
 
@@
 
-\section{parralelizing benchmark and critical analysis}
+\section{parallelising benchmark and critical analysis}
 
@@
 
-\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
 \blindtext % Dummy text

### 2022-12-04 report/article_2.tex
added benchmarks and some analysis

@@
 
-\usepackage[hmarginratio=5:5,top=32mm,columnsep=20pt]{geometry} % Document margins
+\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt,a4paper,total={6in,9in}]{geometry} % Document margins
 \usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
@@
 \usepackage{hyperref} % For hyperlinks in the PDF
-
+\usepackage{graphicx}
+\graphicspath{{./img/}}
 %----------------------------------------------------------------------------------------
@@
 %----------------------------------------------------------------------------------------
-\section{Introduction to the parallelised solution}
+\section{Introduction to the parallel solution}
 
-As an initial attempt, we began by implementing a single-threaded Game of Life (GoL) Engine, expanding upon this work with parallelisation by virtue of concurrent go-routine workers.
-Recognising that branch instructions throttled efficiency of \texttt{getNeighbourCount}, we removed if-statements, introduced bit-masking and assignment which drastically improving runtime.
+As an initial attempt, we began by implementing a 
+single-threaded Game of Life (GoL) Engine, expanding upon this work 
+with parallelisation by virtue of concurrent go-routine workers. 
+Recognising that there exists various methods to calculating the perimeter
+for \texttt{getNeighbourCount}, we tried various counting methods - taking into 
+account the wrap-around feature of Game of Life.
 
-\begin{itemize}
-  \item single threaded implementation
-  \item events passing (alive cells, turn completed)
-  \item PGM image output implementation
-  \item SDL implementation (flipped cells events)
-  \item using channels and goroutines to create multiple workers (parallelising the solution)
-  \item switching to a memory sharing solution instead of using channels (TODO?)
-  \item implementing halo exchange
-\end{itemize}
+% \begin{itemize}
+%   \item single threaded implementation
+%   \item events passing (alive cells, turn completed)
+%   \item PGM image output implementation
+%   \item SDL implementation (flipped cells events)
+%   \item using channels and goroutines to create multiple workers (parallelising the solution)
+%   \item switching to a memory sharing solution instead of using channels (TODO?)
+%   \item implementing halo exchange
+% \end{itemize}
 %------------------------------------------------
@@
 \section{benchmarking methodology}
+Our means of evaluating parallelisation performance gains 
+was via the Go testing framework. Using 512px x 512px PGM images, we 
+repeated each benchmark 10 times, taking the average runtime for each thread to acquire the central value, 
+and plotting them. Each feature implementation was benchmarked in the same fashion.
+% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
+% \begin{itemize}
+%   \item use built in golang benchmark framework
+%   \item run command ....
+%   \item output to txt
+%   \item run python script to plot graphs
+%   \item barchart of time taken against threads used
+%   \item runs benchmarks on all pgm images
+%   \item repeats benchmarks until results fall within a certain standard deviation (TBC)
+% \end{itemize}
+%------------------------------------------------
+\begin{figure}
+  \includegraphics[width=\linewidth]{modulo.png}
+  \caption{Modulo Implementation}
+  \label{fig:chart1}
+\end{figure}
 
-% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-\begin{itemize}
-  \item use built in golang benchmark framework
-  \item run command ....
-  \item output to txt
-  \item run python script to plot graphs
-  \item barchart of time taken against threads used
-  \item runs benchmarks on all pgm images
-  \item repeats benchmarks until results fall within a certain standard deviation (TBC)
-\end{itemize}
-%------------------------------------------------
+\section{parallelising the program}
+The baseline implementation's distributor function starts with
+reading the image into 2D slice via \textit{io} commands, before evenly dividing 
+the workload between workers. With the aid of go-routines, workers have access to the old slice
+as well as the boundaries in which to process. By using \texttt{getNeighbourCount}, we apply the Game of Life
+laws onto the world, piping results into a blank slice of size: \[( endRow - startRow ) \times ImageWidth\]
+Sending the results back to the distributor, we integrate back together the new slices, move the pointer to the old world
+to this newly-merged world and send the appropriate signals on the \textit{events}, \textit{io} channels.
 
-\section{parralelising the program}
-
+\begin{figure}
+  \includegraphics[width=\linewidth]{branching.png}
+  \caption{Branching Implementation}
+  \label{fig:chart2}
+\end{figure}
+\begin{figure}
+  \includegraphics[width=\linewidth]{bitmask.png}
+  \caption{Bitmask Implementation}
+  \label{fig:chart3}
+\end{figure}
 % \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
@@
 
-\section{parallelising benchmark and critical analysis}
+\section{On the topic of analysis}
 
+From Fig. 1, we can conclude that multithreading have an overall improvement on performance.
+Statistically-speaking, we find the solution using 
+modulo operations is most affected by thread usage, having standard deviation 413ms, as opposed to 107ms 
+and 73ms for branching and bitmasking respectively. Interested persons will find our results achieved by:
+\[SD(\forall x \in runtime : max(runtime) - x)\]
+This, however, implies not that higher threads means greater 
+performance as there exists other, non-parallelisable tasks, such as reading 
+PGM images set a base level of execution time. 
 % \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.

### 2022-12-04 report/article_2.tex
added mem-sharing and counting in report

@@
 \fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Running title $\bullet$ May 2016 $\bullet$ Vol. XXI, No. 1} % Custom header text
+\fancyhead[C]{Game of Life  $\bullet$ Computer Systems A $\bullet$ December 2022} % Custom header text
 \fancyfoot[RO,LE]{\thepage} % Custom footer text
@@
 \graphicspath{{./img/}}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\usepackage{txfonts}
+
 %----------------------------------------------------------------------------------------
@@
 repeated each benchmark 10 times, taking the average runtime for each thread to acquire the central value, 
-and plotting them. Each feature implementation was benchmarked in the same fashion.
+and plotting them. Each feature implementation was benchmarked in the same fashion, using Linux lab machines 
+with CPUs being Intel i7-8700 (6 core, 12 threads).
 % \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
@@
 %------------------------------------------------
+\renewcommand\thesubsection{\Alph{subsection}}
 \begin{figure}
@@
 
-\section{parallelising the program}
+\section{A split personality program}
+\subsection{Parallelisation logic}
 The baseline implementation's distributor function starts with
@@
 
+%------------------------------------------------
+
+\subsection{On the topic of analysis 1}
+
+From Fig. 1, we can conclude that multithreading have an overall improvement on performance. 
+There is a correlation between increasing number of threads and decreasing execution times, likely due 
+to some functions, such as \texttt{getNeighbourCount}, being "embarrassingly parallelisable". We 
+favour the term "delightfully parallel". As we have more cores for processing, the time taken for each turn is 
+reduced, hence the curve.
+
+We do take notice of the obvious trend of diminishing returns - 
+the gradient of the time-thread curve flattens. We attribute this to unparallelisable core tasks such as IO 
+- the major slowdown that is reading and writing to/from the disk, and submit that using more threads than 
+a CPU physically has causes overhead in scheduling, minimising the benefits of allocating extra threads for processing.
+We suspect also that, contributing to this, the workload division does not change as much the more threads there are,
+which explains the reciprocal graph shape curve.
+
+Regarding the numerical results, we find greatest performance was with 16 threads (despite the diminishing returns)
+with a 67.6\% decrease in processing time. As for the worst change in performance thread-to-thread, 
+we see the 14th - 15th threads having the smallest difference in performance change (0.353\%), but acknowledge the
+potential anomalous data point.
+
+\section{How to count?}
+\subsection{Counting methods}
+It is well documented that division and other non-atomic arithmetic operations are slow. Thus, we submit our alternative
+implementations for \texttt{getNeighbourCount}. Fig. 1, our original implementation, uses the modulo operation (\%), 
+whereas Fig. 2, Fig. 3, use Branching (if-statements) and Bitmasking respectively. From this data, we see
+that Bitmasking is the fastest on average, but we can identify positive notes from all three.
+\subsection{On the topic of Analysis 2}
 \begin{figure}
@@
 \end{figure}
-% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-\begin{itemize}
-  \item show workers pseudocode
-  \item show handler pseudocode
-  \item flowchart/ diagram of how channel communication workers
-  \item simple flowchart of how the whole program works currently
-\end{itemize}
-%------------------------------------------------
 
-\section{On the topic of analysis}
-
-From Fig. 1, we can conclude that multithreading have an overall improvement on performance.
 Statistically-speaking, we find the solution using 
@@
 and 73ms for branching and bitmasking respectively. Interested persons will find our results achieved by:
-\[SD(\forall x \in runtime : max(runtime) - x)\]
-This, however, implies not that higher threads means greater 
-performance as there exists other, non-parallelisable tasks, such as reading 
-PGM images set a base level of execution time. 
-% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-\begin{itemize}
-  \item show graph for 5120x5120.gpm for single implementation vs parralel implementation
-  \item describe the differences between each graph
-  \item explain the differences by referencing the flowchart/ how the code worker
-  \item explain the shape of the graph (why it resembles a 1/x graph). e.g. bottlenecks in channel communication...
-\end{itemize}
+\[SD(\forall x \in runtimes : max(runtimes) - x)\]
+As discussed before, this, however, implies not that higher threads means greater 
+performance as there exists other, non-parallelisable tasks setting a base level of execution time. 
+The low standard deviation of the bitmasked implementation alludes to this portion of the program being as 
+near optimised as we possibly can.
+
+When it comes to average times for threads, bitmasking takes the lead with
+890ms, followed by 1026ms and 1989ms for branching and modular respectively. This does identify branching for the 
+greatest decrease in runtime, with a 74.2\% decrease, as opposed to 32.1\%.
+
+\subsection{Our Algorithms}
+
+Code-wise, our initial implementation for \texttt{getNeighbourCount} used modulo operators unnecessarily. It did the following:
+\begin{enumerate}[noitemsep]
+  \item Set row = coordinate row \% ImageHeight
+  \item Set col = coordinate col \% ImageWidth
+  \item If row | col < 0, then add ImageHeight/Width respectively
+  \item If world[row][col] was 255, increment alive count
+\end{enumerate}
+
+With this, while it succeeds in its purpose, it does not
+succeed efficiently. Areas of improvement include removal of the modulo operator, which, here,
+does not serve any real purpose. This leads to our second implementation - branching:
+
+\begin{algorithm}
+  \caption{Branching getNeighbourCount}
+  \begin{algorithmic}
+    \State Let $i, j \gets (row, column)$ on world
+  \State Let $O \gets [\{-1,0,1\} \times \{-1, 0, 1\} - \{0,0\}]$
+  \State Let $alive \gets 0$   
+  \For{$(x,y) \in O$}
+      \State $row \gets (i + y)$
+      \If{row < 0}
+        \State $row \gets ImageHeight-1$
+      \ElsIf{row = ImageHeight}
+        \State $row \gets 0$
+      \EndIf
+      \State $col \gets (j + x)$
+      \If{col < 0}
+        \State $col \gets ImageWidth - 1$
+        \ElsIf{col = ImageWidth}
+          \State $col \gets 0$
+      \EndIf
+      \If{world[row][col] = 255}
+      \State $alive++$
+      \EndIf
+    \EndFor
+    \State \Return $alive$
+  \end{algorithmic}
+\end{algorithm}
+Although the line count has certainly increased, the performance has drastically improved. Citing the 
+74.2\% decrease, this does appear to be "good enough" as an implementation. An improvement may be to rid 
+ourselves of branches, as this involves branch prediction which itself is intensive, due to how the CPU 
+does not know which path will be taken ahead of time. If we use only logical operations,
+we get the Bitmasking implementation:
+\begin{algorithm}
+  \caption{Bitmasked getNeighbourCount}
+  \begin{algorithmic}
+    \State Let $i, j \gets (row, column)$ on world
+  \State Let $O \gets [\{-1,0,1\} \times \{-1, 0, 1\} - \{0,0\}]$
+  \State Let $alive \gets 0$   
+  \For{$(x,y) \in O$}
+      \State $row \gets (i + y) \& (ImageHeight - 1)$
+      \State $col \gets (j + x) \& (ImageWidth - 1)$
+      \State $alive \gets alive + world[row][col] >> 7$
+    \EndFor
+    \State \Return alive
+  \end{algorithmic}
+\end{algorithm}
+
+This algorithm is our fastest implementation. Minimising both line count and runtime, it 
+counts the neighbours using purely bitwise operators and assignments - no branching needed.
+
 %------------------------------------------------
@@
 \section{memory sharing}
-
-% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-THIS SECTION WILL BE USED IF WE DECIDE TO IMPLEMENT MEMORY SHARING OVER CHANNEL COMMUNICATION
-%------------------------------------------------
-
-\section{benchmarking and critical analysis of the memory sharing}
+\subsection{Motivation}
+Looking at our achievements in counting, we noticed overhead associated with channel communication.
+Exploring \href{https://github.com/golang/go/blob/4fc9565ffce91c4299903f7c17a275f0786734a1/src/runtime/chan.go}
+{Go's implementation for channels}, we see much overhead. Since we know a channel is a thread-safe memory sharing
+mechanism, we can rid the overhead, using only mutex locks and condition variables.
+\subsection{Implementation}
+Our homemade channel implementation does the following:
+\begin{enumerate}[noitemsep]
+  \item Acquire the mutex lock
+  \item While \texttt{condition} is false, await broadcast unless non-blocking send
+  \item Use the value, then broadcast to other processes
+  \item Finally, unlock mutex
+\end{enumerate}
+We introduce 2 methods, \texttt{Send} \& \texttt{Receive}. \texttt{Send} waits for the memory space 
+to be empty, and replaces it with a value. \texttt{Receive} waits for the memory space to point to a value, and then 
+replaces it with \texttt{nil}. \texttt{Go 1.12} does not have generics, which vastly complicated implementation. 
+Benchmarks with memory sharing (and bitmasked \texttt{getNeighbourCount}) is shown in Fig. 4.
+\subsection{On the topic of Analysis 3}
+benchmarks go here, analysis here
 

### 2022-12-05 report/article_2.tex
finished parallel report

@@
 \end{enumerate}
+\begin{figure}
+  \includegraphics[width=\linewidth]{memshare.png}
+  \caption{Memory Sharing}
+  \label{fig:chart4}
+\end{figure}
 We introduce 2 methods, \texttt{Send} \& \texttt{Receive}. \texttt{Send} waits for the memory space 
@@
 \subsection{On the topic of Analysis 3}
-benchmarks go here, analysis here
+This data was not very analysable. The lack of clear trend and sporadic data points suggests no real difference 
+in performance when adding memory sharing, or faulty implementation (which we realistically do consider a possibility).
+A trend, if any, we identify is that the Modulo performance suffered, but due to the lack of clarity, this is inconclusive.
+We add that repeated trials yielded similar style results. Our method used the following: \[\forall t \in T: ORIG[t] - MEM[t]\]
+Where $MEM$ is memory sharing implementation time, $ORIG$ is channel impl. time, $T$ is the number of threads.
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-\blindtext % Dummy text
+\section{Distributed Implementation}
 
@@
 \section{benchmarking and critical analysis of enabling SDL}
-
+% i dont think we need benchmarking of SDL
 % \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.

### 2022-12-05 report/article_2.tex
added single worker v single thread report

@@
 for \texttt{getNeighbourCount}, we tried various counting methods - taking into 
-account the wrap-around feature of Game of Life.
+account the modularity of Game of Life - as well as finding channel alternatives.
 
@@
 \section{Distributed Implementation}
+Similar to how we did the parallel side, we started the distributed section by using a single worker, this time
+with the worker being a separate process and communicated to via Remote Procedure Calls (RPC henceforth). This expanded to
+doing using multiple workers, implementing fault tolerance and finally halo exchange.
 
 %------------------------------------------------
+\section{A lonesome worker}
+\subsection{Logic}
+Our single-worker distributed implementation moved much of the processing to said worker. From the distributor, we sent a 
+single RPC call to the worker, processed GoL turns on the worker, then sent the final state back. Results are shown in Fig. 5.
+\begin{figure}
+  \includegraphics[width=\linewidth]{para-v-dist.png}
+  \caption{1 worker vs. 1 thread}
+  \label{fig:chart5}
+\end{figure}
+\subsection{On the topic of Analysis 4}
+The differences between the benchmark times is not significant when compared to multi-threaded work. We 
+identified an average of 4.54\% time decrease, which is most likely attributed to how the single-threaded workload had 
+go-routines for Alive Cells and Key presses, whereas the single worker was Step 1 of Distributed, being "barebones".
 

### 2022-12-06 report/article_2.tex
Added fault tol report

@@
 
+\section{Friends? - Multi-worker GoL}
+\subsection{Implementing multiple workers}
+With the introduction of multiple workers, we needed to introduce a broker to interface the workers for the controller
+and vice versa. This did require moving turn iteration logic to the broker, requiring a major refactor. In the end, the system
+revolved around the broker with bidirectional RPC communication betweeen workers and the controller. We equally divide work
+between workers, and all processes keep an internal state of some sort. By massively simplifying, the broker does the following:
+
+\begin{enumerate}[noitemsep]
+  \item Listen for worker and controller connections
+  \item On connect, call the worker/controller back (bidirectional)
+  \item Listen for GoL process request from controller
+  \item Iterate, dividing and sending work to each worker.
+  \item After getting all parts back together, apply changes to broker internal state and divide new state up
+\end{enumerate}
+%--------------
+\begin{itemize}
+  \item ADD BENCHMARKS HERE
+\end{itemize}
+%--------------
+\subsection{Analysis 5}
+%--------------
+\begin{itemize}
+  \item ADD ANALYSIS HERE
+  \item explain trends, averages, maybe std dev?
+  \item if low std dev then time not affected too much by worker count
+\end{itemize}
+%--------------
+\section{Tolerating faults}
+\subsection{Managing workers}
+In our previous implementation, we never checked for worker disconnects. This leaves us in a vulnerable
+position - if a worker were to be terminated at any point, then we would have one section of changes missing.
+This would show itself with a static portion on the SDL window. There is also the case of adding more workers. Either
+way, it means we need to keep the order of workers. We present Algorithm 3 to solve this.
+\begin{algorithm}
+  \caption{Worker-diff checker}
+  \begin{algorithmic}
+    \State Require $ids \subseteq \mathbb{N}$
+    \State Require $oldIds \subseteq \mathbb{N}$ 
+    \State Let $idMap \gets Map$
+    \For{$id \in ids$}
+      \State $idMap[id] \gets addr$
+    \EndFor
+  \For{$id \in oldIds$}
+      \If{idMap[id] = nil}
+      \State \Return False
+      \EndIf
+    \EndFor
+    \State \Return True
+  \end{algorithmic}
+\end{algorithm}
+
+What this algorithm achieves is detection of disconnects and reconnects.
+Given that we have the loop invariant (in turn iterator) "no two workers will have the same worker id",
+we can use this algorithm, and if we detect a change, recalibrate the workers to adjust for new/missing worker. 
+One may ask, does a worker disconnect not throw errors when calling RPC methods? In this case, we utilise \texttt{client.Go},
+which notifies the calling process of errors without crashing. So, if error, then trigger recalibration.
+
 \section{Implementing Halo Exchange}
@@
 TO BE IMPLEMENTED
-%------------------------------------------------
-
-\section{benchmarking and critical analysis of enabling SDL}
-% i dont think we need benchmarking of SDL
-% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-\begin{itemize}
-  \item show graphs and describe shape and differences of with vs without SDL
-  \item explain how SDL works
-  \item explain how this causes bottlenecks
-\end{itemize}
 %------------------------------------------------

### 2022-12-06 worker/worker.go
fixed worker IP

@@ -141,6 +141,7 @@ func (w *Worker) PushHalo(req stubs.PushHaloRequest, res *stubs.NilResponse) (er
 
 func main() {
 	bAddr := flag.String("brokerIP", "127.0.0.1:9000", "IP address of broker")
+	wAddr := flag.String("workerIP", "127.0.0.1", "IP address of broker")
 	pAddr := flag.String("port", "9010", "Port to listen on")
 	flag.Parse()
 
@@ -165,7 +166,7 @@ func main() {
 	worker.broker = broker
 	res := new(stubs.ConnectResponse)
 	req := stubs.ConnectRequest{
-		IP: "127.0.0.1:" + *pAddr,
+		IP: *wAddr + ":" + *pAddr,
 	}
 	broker.Call(stubs.WorkerConnect, req, res)
 	worker.id = res.Id

### 2022-12-06 bench_test.go
something

@@ -11,7 +11,7 @@ func BenchmarkLocal(b *testing.B) {
 	turns := 100
 	threadConfs := []int{1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16}
 	imageConfs := []int{16, 64, 128, 256, 512}
-
+	fmt.Println("hi")
 	for _, imageSize := range imageConfs {
 		for _, threads := range threadConfs {
 			p := gol.Params{

### 2022-12-06 gol/distributor.go

@@ -192,28 +192,19 @@ func distributor(p Params, d distributorChannels, kp <-chan rune) {
 
 	}
 	// exit the broker
-	fmt.Println("0")
 	client.Call(stubs.ControllerQuit, stubs.NilRequest{}, new(stubs.NilResponse))
-	fmt.Println("1")
 	client.Close()
-
-	fmt.Println("2")
 	close(done)
-
-	fmt.Println("3")
 	controller.sendExitSignals()
-	fmt.Println("4")
 	controller.shutDownSequence()
 	fmt.Printf("after\n")
 }
 
 func (c *Controller) shutDownSequence() {
 	// Get the final state of the world
-	fmt.Println("5")
 	world, turn := controller.acknowledgedCells.Get()
 
 	// Output the final image
-	fmt.Println("6")
 	controller.eventsSender.SendOutputPGM(world, turn)
 
 	// TODO: Report the final state using FinalTurnCompleteEvent.
@@ -236,7 +227,6 @@ func (c *Controller) shutDownSequence() {
 // for the user to view in SDL window
 func (c *Controller) PushState(req stubs.PushStateRequest, res *stubs.NilResponse) (err error) {
 	c.m.Lock()
-	fmt.Printf("turn: %d\n", req.Turn)
 	// util.VisualiseSquare(c.acknowledgedCells.CurrentWorld, len(c.acknowledgedCells.CurrentWorld), len(c.acknowledgedCells.CurrentWorld))
 	c.acknowledgedCells.UpdateWorldAndTurn(req.FlippedCells, req.Turn)
 	// util.VisualiseSquare(c.acknowledgedCells.CurrentWorld, len(c.acknowledgedCells.CurrentWorld), len(c.acknowledgedCells.CurrentWorld))

### 2022-12-06 main.go

@@ -17,19 +17,19 @@ func main() {
 	flag.StringVar(
 		&params.ListenIP,
 		"listen",
-		"127.0.0.1",
+		"54.145.138.74",
 		"Specify the IP address of the distributor (this machine). Defaults to 127.0.0.1.")
 
 	flag.IntVar(
 		&params.ListenPort,
 		"p",
-		8090,
+		8030,
 		"Specify the port to listen on. Defaults to 8090.")
 
 	flag.StringVar(
 		&params.BrokerAddr,
 		"broker",
-		"127.0.0.1:9000",
+		"100.26.97.47:8030",
 		"Specify the socket address of the broker. Defaults to 127.0.0.1:9000.")
 
 	flag.IntVar(
@@ -58,7 +58,7 @@ func main() {
 
 	noVis := flag.Bool(
 		"noVis",
-		false,
+		true,
 		"Disables the SDL window, so there is no visualisation during the tests.")
 
 	flag.Parse()

### 2022-12-06 broker/BrokerLib.go
progress quit

@@ -154,7 +154,6 @@ func (b *Broker) getWorkersNotResponded() []int {
 // IMPORTANT: must mutex lock in calling scope
 func (b *Broker) primeWorkers(firstTime bool) {
 	b.Mu.Lock()
-	defer b.Mu.Unlock()
 	divideEvenly(b)
 	currRowOffset := 0
 	//this is assuming that the order of the map does not change within this function
@@ -175,12 +174,17 @@ func (b *Broker) primeWorkers(firstTime bool) {
 			FirstTime:   firstTime,
 		}
 		worker := b.Workers[id].client
+		println("before call")
+		b.Mu.Unlock()
 		err := worker.Call(stubs.InitialiseWorker, initRequest, new(stubs.NilResponse))
+		println("after call")
 		if err != nil {
 			fmt.Printf("err: %s\n", err)
 		}
 		currRowOffset += workSize //keeping track of currentRowOffset
+		b.Mu.Lock()
 	}
+	b.Mu.Unlock()
 	fmt.Println("Finished priming")
 }
 

### 2022-12-06 broker/broker.go

@@ -72,6 +72,7 @@ func (b *Broker) StartGOL(req stubs.StartGOLRequest, res *stubs.NilResponse) (er
 		// fmt.Printf("lastCompletedTurn: %d totalTurns: %d controllerNil?: %t\n", b.lastCompletedTurn, req.P.Turns, b.Controller == nil)
 		select {
 		case badWorkerId := <-b.errorChan: //if error: restart workers
+			fmt.Printf("R\n")
 			// reset all vars
 			//remove bad worker with id
 			b.resetBroker(badWorkerId)
@@ -80,6 +81,7 @@ func (b *Broker) StartGOL(req stubs.StartGOLRequest, res *stubs.NilResponse) (er
 			lastPushState = time.Now()
 
 		case <-b.processCellsReq:
+			fmt.Printf("A\n")
 			b.Mu.Lock()
 			if b.Controller == nil {
 				b.Mu.Unlock()
@@ -102,6 +104,8 @@ func (b *Broker) StartGOL(req stubs.StartGOLRequest, res *stubs.NilResponse) (er
 			lastPushState = time.Now()
 			b.Mu.Unlock()
 		default:
+			fmt.Printf("B\n")
+			time.Sleep(time.Millisecond * 100)
 			if time.Now().Second()-lastPushState.Second() > int(time.Second.Seconds()) {
 				if b.Controller == nil {
 					lastPushState = time.Now()
@@ -116,8 +120,10 @@ func (b *Broker) StartGOL(req stubs.StartGOLRequest, res *stubs.NilResponse) (er
 		}
 
 	}
-	b.resetBroker(-1)
+	fmt.Printf("E\n")
 	b.primeWorkers(false) //to stop the workers from carrying on executing turns when controller disconnects before all turns have been processed
+	fmt.Printf("F\n")
+	b.resetBroker(-1)
 	fmt.Println("brokerStartGol Done")
 	return
 }
@@ -150,11 +156,12 @@ func (b *Broker) PauseState(req stubs.NilRequest, res *stubs.PauseResponse) (err
 }
 
 func (b *Broker) PushState(req stubs.BrokerPushStateRequest, res *stubs.NilResponse) (err error) {
-	b.Mu.Lock()
-	defer b.Mu.Unlock()
+	fmt.Printf("PushState\n")
 	if b.exit || b.Controller == nil {
 		return
 	}
+	b.Mu.Lock()
+	defer b.Mu.Unlock()
[…]

### 2022-12-06 gol/distributor.go

@@ -41,14 +41,14 @@ var controller Controller
 func (c *Controller) aliveCellsTicker(client *rpc.Client) {
 	ticker := time.NewTicker(2 * time.Second)
 	defer ticker.Stop()
-
-	for {
+	exit := false
+	for !exit {
 		select {
 		case <-ticker.C:
 			aliveCellsCount, turn := c.acknowledgedCells.GetAliveCount()
 			c.eventsSender.SendAliveCellsList(turn, aliveCellsCount)
 		case <-c.exitChannels[aliveCellsTickerFunc]:
-			break
+			exit = true
 		}
 	}
 
@@ -112,13 +112,11 @@ func newController() Controller {
 }
 
 func (c *Controller) sendExitSignals() {
-	c.m.Lock()
-	go func() {
-		c.exitChannels[aliveCellsTickerFunc] <- true
-	}()
+	// go func() {
+	c.exitChannels[aliveCellsTickerFunc] <- true
+	// }()
 	c.exitChannels[receiver] <- true
 
-	c.m.Unlock()
 }
 
 // distributor distributes the work to the broker via rpc calls

### 2022-12-06 worker/worker.go

@@ -39,8 +39,6 @@ func (w *Worker) EvolveSlice(req stubs.WorkRequest, res *stubs.NilResponse) (err
 	topDone <- new(rpc.Call)
 	botDone := make(chan *rpc.Call, 1)
 	botDone <- new(rpc.Call)
-	// brokerDone := make(chan *rpc.Call, 1)
-	// wg := new(sync.WaitGroup)
 	w.repriming = false
 	w.inEvolveLoop = true
 	for i := req.StartTurn; (i <= req.Turns || req.Turns == -1) && !w.repriming; i++ { //if -1 that means forever
@@ -51,34 +49,49 @@ func (w *Worker) EvolveSlice(req stubs.WorkRequest, res *stubs.NilResponse) (err
 		}
 		var evolvedSlice [][]byte = createNewSlice(w.height, w.width) // TODO try move this outside of for loop and see if it still works
 		// wait for Halo input
-		var topHalo []byte
-		var botHalo []byte
+		var topHalo []byte = nil
+		var botHalo []byte = nil
+		fmt.Printf("A\n")
 		if req.IsSingleWorker { //if single worker
 			topHalo = w.container.CurrentWorld[w.height-1] //set to last row
 			botHalo = w.container.CurrentWorld[0]          //set to first row
 		} else { //if more than 1 worker
-			topHalo = <-w.topHalo
-			botHalo = <-w.botHalo
+			exit := false
+			for (topHalo == nil || botHalo == nil) && !exit {
+				select {
+				case topHalo = <-w.topHalo:
+
+				case botHalo = <-w.botHalo:
+				default:
+					if w.repriming {
+						fmt.Println("REPRIMING IS TRUE")
+						exit = true
+					}
+				}
+			}
+
 		}
+		if w.repriming {
+			continue
+		}
+		fmt.Printf("B\n")
 		// perform iteration
 		flipped := w.evolve(evolvedSlice, topHalo, botHalo)
+		fmt.Printf("C\n")
 		// updated world in worker api
 		w.container.UpdateWorld(evolvedSlice)
 
-		// TODO pushFLippedCells to broker
+		fmt.Printf("D\n")
+		// pushFLippedCells to broker
 		brokerReq := stubs.BrokerPushStateRequest{
 			FlippedCells: flipped,
 			Turn:         i,
 			WorkerId:     w.id,
 		}
[…]

### 2022-12-06 report/article_2.tex
fixing Figure 5

@@
 
-\subsection{On the topic of analysis 1}
+\subsection{On the topic of Analysis 1}
 
@@
 \begin{figure}
-  \includegraphics[width=\linewidth]{para-v-dist.png}
+  \includegraphics[width=\linewidth]{1v1.png}
   \caption{1 worker vs. 1 thread}
@@
 \subsection{On the topic of Analysis 4}
-The differences between the benchmark times is not significant when compared to multi-threaded work. We 
-identified an average of 4.54\% time decrease, which is most likely attributed to how the single-threaded workload had 
-go-routines for Alive Cells and Key presses, whereas the single worker was Step 1 of Distributed, being "barebones".
+The absurd differences we observed in the 1 worker vs. 1 thread (distributed versus parallel), is due 
+to a number of things:
+\begin{enumerate}
+  \item Broker and Workers were positioned in US-east-1, meaning North Virginia, which is 7000km away.
+  This oversight caused the massive difference in results.
+  \item RPC calls for 1 worker may have caused overhead (extremely minor compared to above)
+  \item Controller was not positioned in US-east-1, as we were not able to successfully run the controller
+  program on the EC2 instance, due to issues with SDL.
+\end{enumerate}
+Due to this, the results are statistically worthless.
 
@@
 %--------------
-\begin{itemize}
-  \item ADD BENCHMARKS HERE
-\end{itemize}
+% \begin{figure}
+%   \includegraphics[width=\linewidth]{multiworker.png}
+%   \caption{Multi-worker performance}
+%   \label{fig:chart6}
+% \end{figure}
 %--------------
@@
 \section{Tolerating faults}
-\subsection{Managing workers}
 In our previous implementation, we never checked for worker disconnects. This leaves us in a vulnerable
@@
 Given that we have the loop invariant (in turn iterator) "no two workers will have the same worker id",
-we can use this algorithm, and if we detect a change, recalibrate the workers to adjust for new/missing worker. 
+we can use this algorithm. If we detect a change, recalibrate the workers to adjust for new/missing worker. 
 One may ask, does a worker disconnect not throw errors when calling RPC methods? In this case, we utilise \texttt{client.Go},
-which notifies the calling process of errors without crashing. So, if error, then trigger recalibration.
+which notifies the calling process of errors without crashing. So, if error, then trigger recalibration by resetting each worker's
+internal state.
 
 \section{Implementing Halo Exchange}
+\subsection{How Jesus Christ came to be}
+Our first attempt wasn't exactly the definition of halo exchange. It used halos distributed by broker to
+each worker, who then used the halos to update their internal state. Now, our implementation transfers the turn iterator
+logic to the workers. The sequence of events happens as follows:
+\begin{enumerate}[noitemsep]
+  \item Worker connects to broker, and is initialised with parameters of controller.
+  \item On initialisation, worker will dial the top and bottom neighbours (if any), then await work.
+  \item On work, send its halos, and evolve its own slice given the neighbouring halos
+  \item Results are pushed to broker and loop
+\end{enumerate}
+This oversimplified view of halo exchange is represented by a significant amount of implementation code.
+We present our results with Fig. 6.
 
-% \lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-TO BE IMPLEMENTED
-%------------------------------------------------
+\subsection{Analysis 6 - Judgement Day}
+An obvious pattern we see is that the time taken is significantly reduced. We find that there is a negative
+correlation between workers and time - increasing workers decreases the runtime of the program. We argue that
+these results stems from the reduced overhead of needing to use the broker to transfer halos; decreased workload
+from the broker having to manage turns actually becoming a broker seen typically in Publish-Subscriber models.
 
-\section{benchmarking and critical analysis of the Halo Exchange}
-
-\lettrine[nindent=0em,lines=3]{L} orem ipsum dolor sit amet, consectetur adipiscing elit.
-TO BE IMPLEMENTED
 %------------------------------------------------
@@
 \section{comparing the parralel benchmarks}
-
-Maecenas sed ultricies felis. Sed imperdiet dictum arcu a egestas. 
-\begin{itemize}
-\item Donec dolor arcu, rutrum id molestie in, viverra sed diam
-\item Curabitur feugiat
-\item turpis sed auctor facilisis
-\item arcu eros accumsan lorem, at posuere mi diam sit amet tortor
-\item Fusce fermentum, mi sit amet euismod rutrum
-\item sem lorem molestie diam, iaculis aliquet sapien tortor non nisi
-\item Pellentesque bibendum pretium aliquet
-\end{itemize}
-\blindtext % Dummy text
-
-Text requiring further explanation\footnote{Example footnote}.
 
@@
 
-\blindtext % Dummy text
-
-\begin{equation}
-\label{eq:emc}
-e = mc^2
-\end{equation}
-
-\blindtext % Dummy text
-
 %------------------------------------------------
@@
 
-A statement requiring citation \cite{Figueredo:2009dg}.
-\blindtext % Dummy text
-
+\begin{itemize}
+  \item Memory sharing implementation for channels - we were able to create an alternative to channels, but
+  this ended up more or less the same performance as a regular channel, if not worse. A strong area of improvement
+  would be to rewrite this, and substitute all the other channels used with our implementation
+  \item Parallelising workers - we did not manage to implement this in time. Since all the code for multithreading
+  was in our parallel code, we could have imported the code then implemented it inside our workers.
+\end{itemize}
 \subsection{Thanks}
-
-\blindtext % Dummy text
+\begin{itemize}
+  \item Lecturers and TAs for their support
+  \item Monster Energy\texttrademark
+  \item My missing sanity
+\end{itemize}
 
@@
 
-\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template
-
-\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
-Figueredo, A.~J. and Wolf, P. S.~A. (2009).
-\newblock Assortative pairing and life history strategy - a cross-cultural
-  study.
-\newblock {\em Human Nature}, 20:317--330.
-
-\end{thebibliography}
-
 %----------------------------------------------------------------------------------------

### 2022-12-06 worker/worker.go
progress

@@ -45,9 +45,11 @@ func (w *Worker) EvolveSlice(req stubs.WorkRequest, res *stubs.NilResponse) (err
 		fmt.Printf("processing turn: %d of turns %d\n", i, req.Turns)
 		// send Halo to adjacent workers
 		if !req.IsSingleWorker {
-			w.pushHalos(topDone, botDone) //maybe check if they have been received before send next?
+			go w.pushHalos(topDone, botDone) //maybe check if they have been received before send next?
 		}
+		fmt.Printf("1\n")
 		var evolvedSlice [][]byte = createNewSlice(w.height, w.width) // TODO try move this outside of for loop and see if it still works
+		fmt.Printf("2\n")
 		// wait for Halo input
 		var topHalo []byte = nil
 		var botHalo []byte = nil
@@ -144,6 +146,19 @@ func (w *Worker) Shutdown(req stubs.NilRequest, res *stubs.NilResponse) (err err
 }
 
 func (w *Worker) PushHalo(req stubs.PushHaloRequest, res *stubs.NilResponse) (err error) {
+	for w.repriming {
+		if req.IsTop {
+			select {
+			case w.topHalo <- req.Halo:
+			default:
+			}
+		} else {
+			select {
+			case w.botHalo <- req.Halo:
+			default:
+			}
+		}
+	}
 	if req.IsTop {
 		w.topHalo <- req.Halo
 	} else {

### 2022-12-06 report/article_2.tex
finalise report?

@@
 \subsection{On the topic of Analysis 4}
-The absurd differences we observed in the 1 worker vs. 1 thread (distributed versus parallel), is due 
-to a number of things:
-\begin{enumerate}
-  \item Broker and Workers were positioned in US-east-1, meaning North Virginia, which is 7000km away.
-  This oversight caused the massive difference in results.
-  \item RPC calls for 1 worker may have caused overhead (extremely minor compared to above)
-  \item Controller was not positioned in US-east-1, as we were not able to successfully run the controller
-  program on the EC2 instance, due to issues with SDL.
-\end{enumerate}
-Due to this, the results are statistically worthless.
+We decided to test single threaded parallel version against a single worker distributed GoL. Seeing that the 1 worker ran with
+runtime 1362ms versus the single-threaded 1208ms, we concluded that the 154ms difference was due to the latency between Bristol
+and AWS London, in addition to some potentially contributing overhead from RPC calls. The rest of the runtime we concluded was the natural
+runtime of the worker processing Game of Life.
 
@@
 \end{enumerate}
+In order to call the worker back, this meant that the worker needed to provide its own IP address - the same 
+process happens for the controller. The reason why the worker needs bidirectional communication is that later on with halo-exchange,
+we will need this to push the state when the turn iterator is at the worker, and also to allow workers to disconnect 
+without errors.
+
+When it comes to the reassembly, the worker sends the flipped cells to the broker, and because no 
+worker interferes with another worker's section, we can just apply the changes in any order to the broker's
+internal state.
 %--------------
-% \begin{figure}
-%   \includegraphics[width=\linewidth]{multiworker.png}
-%   \caption{Multi-worker performance}
-%   \label{fig:chart6}
-% \end{figure}
+\begin{figure}
+  \includegraphics[width=\linewidth]{multi-worker.png}
+  \caption{Multi-worker local performance}
+  \label{fig:chart6}
+\end{figure}
 %--------------
@@
 %--------------
-\begin{itemize}
-  \item ADD ANALYSIS HERE
-  \item explain trends, averages, maybe std dev?
-  \item if low std dev then time not affected too much by worker count
-\end{itemize}
-%--------------
+Shown in Fig. 6 is the distributed system when run on a local machine (due to port-forwarding issues).
+We can see a general negative correlation between workers and runtime - as workers increases, the runtime decreases.
+Noticing that the runtime begins to incline at the end, we initially thought this was due to RPC overhead, but it may have been
+due to how running on 1 machine starts to cost more performance than it improves.
+
 \section{Tolerating faults}
@@
 
+More specifically, the broker sends an RPC call to each remaining worker with a fresh state from which to 
+evolve from. The previous internal state of each worker is discarded, even if it is one turn ahead as we use the broker's
+internal state, which is only pushed to if all workers return their flipped cells fine.
+\begin{figure}
+  \includegraphics[width=\linewidth]{halo.png}
+  \caption{Multi-worker AWS performance}
+  \label{fig:chart7}
+\end{figure}
+
 \section{Implementing Halo Exchange}
@@
 \subsection{Analysis 6 - Judgement Day}
-An obvious pattern we see is that the time taken is significantly reduced. We find that there is a negative
-correlation between workers and time - increasing workers decreases the runtime of the program. We argue that
-these results stems from the reduced overhead of needing to use the broker to transfer halos; decreased workload
-from the broker having to manage turns actually becoming a broker seen typically in Publish-Subscriber models.
+The absurd difference in value between Fig. 6, Fig. 7, was due to what we identified as transmission latency. The broker
+was positioned in \texttt{us-east-1}, North Virginia, and the 7000km+ differential caused the latency 
+to be very high.
+We did notice, despite the glaringly obvious latency issue, that the time did start to decline as more workers
+were added, implying that there may have been some speed increases due to our distributed system. However,
+we do not reject the notion of these being oscillations in latency values.
+
+Giving the benefit of the doubt, had the EC2 instances been set up in the correct region, then we should have noticed
+a significantly smaller runtime with a lower standard deviation - that is, less variance in data points.
+%------------------------------------------------
 
@@
 
-\section{comparing the parralel benchmarks}
-
-%------------------------------------------------
-
-\section{Results}
-
-\begin{table}
-\caption{Example table}
-\centering
-\begin{tabular}{llr}
-\toprule
-\multicolumn{2}{c}{Name} \\
-\cmidrule(r){1-2}
-First name & Last Name & Grade \\
-\midrule
-John & Doe & $7.5$ \\
-Richard & Miles & $2$ \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-%------------------------------------------------
-
-\section{Discussion}
-
-\subsection{Areas for improvement}
+\section{Areas for improvement}
 
@@
   was in our parallel code, we could have imported the code then implemented it inside our workers.
+  \item Controller on AWS node - we found issues with latency when using AWS nodes, so if the "local" controller was
+  localised in the AWS data center, then we would eliminate this issue. We, however, ran into issues with SDL not working
+  on EC2 machines and port-forward which meant these latency issues still happened.
+  \item Broker fault tolerance - As a extreme extension, we have the broker as our single point of failure. If the broker fails, then
+  nothing can connect and no processes can occur. We could, in theory, create a second broker to be redundant in case the instance 
+  running the initial broker fails.
 \end{itemize}

### 2022-12-06 bench_test.go
implemented pause

@@ -9,8 +9,8 @@ import (
 
 func BenchmarkLocal(b *testing.B) {
 	turns := 100
-	threadConfs := []int{1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16}
-	imageConfs := []int{16, 64, 128, 256, 512}
+	threadConfs := []int{1}
+	imageConfs := []int{512}
 
 	for _, imageSize := range imageConfs {
 		for _, threads := range threadConfs {

### 2022-12-06 broker/broker.go

@@ -117,6 +117,7 @@ func (b *Broker) StartGOL(req stubs.StartGOLRequest, res *stubs.NilResponse) (er
 					b.errorChan <- id
 				}
 			}
+			b.Pause.Wait()
 		}
 
 	}
@@ -144,7 +145,7 @@ func (b *Broker) ControllerQuit(req stubs.NilRequest, res *stubs.NilResponse) (e
 }
 
 // pauses the game of life loop
-func (b *Broker) PauseState(req stubs.NilRequest, res *stubs.PauseResponse) (err error) {
+func (b *Broker) PauseState(req stubs.NilRequest, res *stubs.NilRequest) (err error) {
 	if b.isPaused {
 		b.Pause.Done()
 		b.isPaused = false
@@ -152,11 +153,16 @@ func (b *Broker) PauseState(req stubs.NilRequest, res *stubs.PauseResponse) (err
 		b.Pause.Add(1)
 		b.isPaused = true
 	}
+
+	for id := range b.workerIds {
+		b.Workers[id].client.Call(stubs.WorkerPauseState, new(stubs.NilRequest), new(stubs.NilResponse))
+	}
 	return
 }
 
 func (b *Broker) PushState(req stubs.BrokerPushStateRequest, res *stubs.NilResponse) (err error) {
 	fmt.Printf("PushState\n")
+	b.Pause.Wait()
 	if b.exit || b.Controller == nil {
 		return
 	}

### 2022-12-06 stubs/stubs.go

@@ -38,6 +38,7 @@ var ControllerQuit = "Broker.ControllerQuit"
 
 // we use this to indicate the controller wants to pause processing
 var PauseState = "Broker.PauseState"
+var WorkerPauseState = "Worker.PauseState"
 
 // req NilRequest, 		sends PauseResponse   x
 // - server must send the turn currently being processed

### 2022-12-06 worker/worker.go

@@ -7,6 +7,7 @@ import (
 	"net/rpc"
 	"os"
 	"os/signal"
+	"sync"
 	"syscall"
 	"time"
 
@@ -29,6 +30,8 @@ type Worker struct {
 	botHalo      chan []byte
 	exit         bool
 	inEvolveLoop bool
+	Pause        sync.WaitGroup
+	isPaused     bool
 }
 
 func (w *Worker) EvolveSlice(req stubs.WorkRequest, res *stubs.NilResponse) (err error) {
@@ -90,6 +93,7 @@ func (w *Worker) EvolveSlice(req stubs.WorkRequest, res *stubs.NilResponse) (err
 			Turn:         i,
 			WorkerId:     w.id,
 		}
+		w.Pause.Wait()
 		w.broker.Call(stubs.BrokerPushState, brokerReq, new(stubs.NilResponse))
 
 		fmt.Printf("E\n")
@@ -137,6 +141,17 @@ func (w *Worker) InitialiseWorker(req stubs.InitWorkerRequest, res *stubs.NilRes
 	return
 }
 
+func (w *Worker) PauseState(req stubs.NilRequest, res *stubs.PauseResponse) (err error) {
+	if w.isPaused {
+		w.Pause.Done()
+		w.isPaused = false
+	} else {
+		w.Pause.Add(1)
+		w.isPaused = true
+	}
+	return
+}
+
 // sent by broker to sleep the distributed system
 func (w *Worker) Shutdown(req stubs.NilRequest, res *stubs.NilResponse) (err error) {
 	// programmatic Ctrl-C
@@ -146,24 +161,30 @@ func (w *Worker) Shutdown(req stubs.NilRequest, res *stubs.NilResponse) (err err
 }
 
 func (w *Worker) PushHalo(req stubs.PushHaloRequest, res *stubs.NilResponse) (err error) {
-	for w.repriming {
-		if req.IsTop {
-			select {
-			case w.topHalo <- req.Halo:
-			default:
-			}
-		} else {
-			select {
-			case w.botHalo <- req.Halo:
-			default:
-			}
-		}
-	}
+	// loop := true
+	fmt.Println("x")
+	// for !w.repriming && loop {
+	// 	if req.IsTop {
+	// 		select {
+	// 		case w.topHalo <- req.Halo:
+	// 			loop = false
+	// 		default:
+	// 		}
+	// 	} else {
+	// 		select {
[…]

### 2022-12-06 report/article_2.tex
temp

@@
   We intend to explain our two solutions for implementing Conway's Game of Life in Golang, namely the Parallel and Distributed versions.
-  Within this report, we will discuss the impacts of thread usage in both our implementations, as well as the effect of distributed computation as opposed to single-machine computation.
+  Within this report, we will discuss the impacts of thread usage in the parallel implementation, as well as the effect of distributed computation as opposed to single-machine computation.
 In essence, we found that execution time was sped up as we increased thread count, reduced branch instructions and implemented a halo exchange mechanism.
@@
 \end{enumerate}
+
+\begin{algorithm}
+  \caption{modulo getNeighbourCount function}
+  \begin{algorithmic}
+    \State Let $row \gets rowCoordinate \% ImageHeight$
+    \State Let $col \gets colCoordinate \% ImageWidth
+    \If{row || col < 0}
+
+    \EndIf
+
+  %   \State Let $i, j \gets (row, column)$ on world
+  % \State Let $O \gets [\{-1,0,1\} \times \{-1, 0, 1\} - \{0,0\}]$
+  % \State Let $alive \gets 0$   
+  % \For{$(x,y) \in O$}
+  %     \State $row \gets (i + y)$
+  %     \If{row < 0}
+  %       \State $row \gets ImageHeight-1$
+  %     \ElsIf{row = ImageHeight}
+  %       \State $row \gets 0$
+  %     \EndIf
+  %     \State $col \gets (j + x)$
+  %     \If{col < 0}
+  %       \State $col \gets ImageWidth - 1$
+  %       \ElsIf{col = ImageWidth}
+  %         \State $col \gets 0$
+  %     \EndIf
+  %     \If{world[row][col] = 255}
+  %     \State $alive++$
+  %     \EndIf
+  %   \EndFor
+  %   \State \Return $alive$
+  \end{algorithmic}
+\end{algorithm}
 

### 2022-12-06 report/article_2.tex
final report

--- 2022-12-06 report/article_2.tex
@@
 \end{enumerate}
+
+\begin{algorithm}
+  \caption{modulo getNeighbourCount function}
+  \begin{algorithmic}
+    \State Let $row \gets rowCoordinate \% ImageHeight$
+    \State Let $col \gets colCoordinate \% ImageWidth$
+    \If{row < 0} 
+    \State $row \gets ImageHeight - 1$
+    \ElsIf{col < 0}
+    \State $row \gets ImageWidth - 1$ 
+    \EndIf
+    \If{alive[row][col] = 0xFF}
+    \State $alive++$
+    \EndIf
+    \State \Return $alive$
+  \end{algorithmic}
+\end{algorithm}
 
@@
 \section{Implementing Halo Exchange}
+Now that we have a broker and a Publish-Subscriber model, we can further optimise by having halo exchange.
 \subsection{How Jesus Christ came to be}

### 2022-12-06 broker/broker.go
removed messages

@@ -11,8 +11,6 @@ import (
 	"uk.ac.bris.cs/gameoflife/stubs"
 )
 
-const doneBuffer = 10
-
 // initialises bidirectional comms with controller
 func (b *Broker) ControllerConnect(req stubs.ConnectRequest, res *stubs.NilResponse) (err error) {
 	fmt.Println("Received controller connect request")
@@ -64,12 +62,9 @@ func (b *Broker) StartGOL(req stubs.StartGOLRequest, res *stubs.NilResponse) (er
 	// start workers
 	b.startWorkers(req)
 	// check for errors
-	// fmt.Printf("lastCompletedTurn: %d totalTurns: %d controllerNil?: %t\n", b.lastCompletedTurn, req.P.Turns, b.Controller == nil)
 	lastPushState := time.Now()
-	// done := make(chan *rpc.Call, doneBuffer)
 	for b.lastCompletedTurn < req.P.Turns && b.Controller != nil {
 
-		// fmt.Printf("lastCompletedTurn: %d totalTurns: %d controllerNil?: %t\n", b.lastCompletedTurn, req.P.Turns, b.Controller == nil)
 		select {
 		case badWorkerId := <-b.errorChan: //if error: restart workers
 			fmt.Printf("R\n")
@@ -89,9 +84,7 @@ func (b *Broker) StartGOL(req stubs.StartGOLRequest, res *stubs.NilResponse) (er
 			}
 			b.lastCompletedTurn++
 			//send error if we don't get here within a second
-			// fmt.Printf("workers responded: %d, turn %d\n", workersRespondedCount, b.lastCompletedTurn)
 			//update controller
-			fmt.Printf("turn %d\n", b.lastCompletedTurn)
 			pushReq := stubs.PushStateRequest{
 				FlippedCells: b.flippedCells[b.lastCompletedTurn],
 				Turn:         b.lastCompletedTurn,
@@ -104,7 +97,6 @@ func (b *Broker) StartGOL(req stubs.StartGOLRequest, res *stubs.NilResponse) (er
 			lastPushState = time.Now()
 			b.Mu.Unlock()
 		default:
-			fmt.Printf("B\n")
 			time.Sleep(time.Millisecond * 100)
 			if time.Now().Second()-lastPushState.Second() > int(time.Second.Seconds()) {
 				if b.Controller == nil {
[…]

### 2022-12-06 gol/distributor.go

@@ -190,55 +190,37 @@ func distributor(p Params, d distributorChannels, kp <-chan rune) {
 
 	}
 	// exit the broker
-	fmt.Println("0")
 	client.Call(stubs.ControllerQuit, stubs.NilRequest{}, new(stubs.NilResponse))
-	fmt.Println("1")
 	client.Close()
 
-	fmt.Println("2")
 	close(done)
 
-	fmt.Println("3")
 	controller.sendExitSignals()
-	fmt.Println("4")
 	controller.shutDownSequence()
-	fmt.Printf("after\n")
 }
 
 func (c *Controller) shutDownSequence() {
 	// Get the final state of the world
-	fmt.Println("5")
 	world, turn := controller.acknowledgedCells.Get()
 
 	// Output the final image
-	fmt.Println("6")
 	controller.eventsSender.SendOutputPGM(world, turn)
 
 	// TODO: Report the final state using FinalTurnCompleteEvent.
-	fmt.Println("7")
 	controller.eventsSender.SendFinalTurn(turn, stubs.GetAliveCells(world)) //turn+1? DEBUG
 	// Make sure that the Io has finished any output before exiting.
-	fmt.Println("8")
 	c.eventsSender.C.ioCommand <- ioCheckIdle
-	fmt.Println("9")
 	<-c.eventsSender.C.ioIdle
-	// fmt.Printf("before\n")
 	c.eventsSender.C.events <- StateChange{turn, Quitting}
 	// Close the channel to stop the SDL goroutine gracefully. Removing may cause deadlock.
-	fmt.Println("10")
 	close(c.eventsSender.C.events)
-	fmt.Println("11")
 }
 
 // This method will be called if the Broker has a calculated new state
 // for the user to view in SDL window
 func (c *Controller) PushState(req stubs.PushStateRequest, res *stubs.NilResponse) (err error) {
 	c.m.Lock()
-	fmt.Printf("turn: %d\n", req.Turn)
-	// util.VisualiseSquare(c.acknowledgedCells.CurrentWorld, len(c.acknowledgedCells.CurrentWorld), len(c.acknowledgedCells.CurrentWorld))
 	c.acknowledgedCells.UpdateWorldAndTurn(req.FlippedCells, req.Turn)
[…]

### 2022-12-06 worker/worker.go

@@ -37,7 +37,6 @@ type Worker struct {
 func (w *Worker) EvolveSlice(req stubs.WorkRequest, res *stubs.NilResponse) (err error) {
 	// the worker has been reprimed, so its internal state is empty
 	w.container.UpdateWorldAndTurn(req.FlippedCells, 0)
-	// TODO: change so that worker keeps track of own turn
 	topDone := make(chan *rpc.Call, 1)
 	topDone <- new(rpc.Call)
 	botDone := make(chan *rpc.Call, 1)
@@ -45,18 +44,14 @@ func (w *Worker) EvolveSlice(req stubs.WorkRequest, res *stubs.NilResponse) (err
 	w.repriming = false
 	w.inEvolveLoop = true
 	for i := req.StartTurn; (i <= req.Turns || req.Turns == -1) && !w.repriming; i++ { //if -1 that means forever
-		fmt.Printf("processing turn: %d of turns %d\n", i, req.Turns)
 		// send Halo to adjacent workers
 		if !req.IsSingleWorker {
 			go w.pushHalos(topDone, botDone) //maybe check if they have been received before send next?
 		}
-		fmt.Printf("1\n")
 		var evolvedSlice [][]byte = createNewSlice(w.height, w.width) // TODO try move this outside of for loop and see if it still works
-		fmt.Printf("2\n")
 		// wait for Halo input
 		var topHalo []byte = nil
 		var botHalo []byte = nil
-		fmt.Printf("A\n")
 		if req.IsSingleWorker { //if single worker
 			topHalo = w.container.CurrentWorld[w.height-1] //set to last row
 			botHalo = w.container.CurrentWorld[0]          //set to first row
@@ -79,14 +74,11 @@ func (w *Worker) EvolveSlice(req stubs.WorkRequest, res *stubs.NilResponse) (err
 		if w.repriming {
 			continue
 		}
-		fmt.Printf("B\n")
 		// perform iteration
 		flipped := w.evolve(evolvedSlice, topHalo, botHalo)
-		fmt.Printf("C\n")
 		// updated world in worker api
 		w.container.UpdateWorld(evolvedSlice)
 
-		fmt.Printf("D\n")
 		// pushFLippedCells to broker
 		brokerReq := stubs.BrokerPushStateRequest{
[…]

### 2022-12-06 worker/worker.go
f

@@ -46,7 +46,7 @@ func (w *Worker) EvolveSlice(req stubs.WorkRequest, res *stubs.NilResponse) (err
 	for i := req.StartTurn; (i <= req.Turns || req.Turns == -1) && !w.repriming; i++ { //if -1 that means forever
 		// send Halo to adjacent workers
 		if !req.IsSingleWorker {
-			go w.pushHalos(topDone, botDone) //maybe check if they have been received before send next?
+			w.pushHalos(topDone, botDone) //maybe check if they have been received before send next?
 		}
 		var evolvedSlice [][]byte = createNewSlice(w.height, w.width) // TODO try move this outside of for loop and see if it still works
 		// wait for Halo input
